### 2020-01-16 ISSTA/experiments.tex
ok

Table \ref{tab:overviewexp} shows a summary of our experimental
results.  In all cases, our approach was able to detect interesting
nondeterminism (in fact, all nondeterminism of which we are aware).  The overhead on testing imposed by nondeterminism
checks varied considerably, but in the worst case did not double the
average cost of each action (the expected overhead, since we
essentially execute each test twice), and thus never halved the amount of
testing performed.  Test reduction always reduced the size of
generated tests by 85\% or more, in less than 15 minutes, thus showing that reduction
has a large payoff at an acceptable price.  Values here and discussed
below are, unless otherwise noted, the mean of 30 runs; we omit variance due to little
variance exhibited, but can provide full details in our replication package.

@@

Our first subject is an implementation of a parallel merge
sort\footnote{\url{https://gist.github.com/stephenmcd/39ded69946155930c347}}. When a sorting
algorithm is implemented using concurrency, a key question to ask is
whether the ordering of (equal under comparison) elements is consistent.
This is related to the question of whether a sort is stable,
but imposes a less strict requirement; a concurrent sort
might not be stable, but might still produce the same (non-stable)
ordering from the same provided sequence, every time.  This is a
particular instance of a common expectation:  while a parallel
implementation of an algorithm will likely have internal
nondeterminism, it is often expected to
behave deterministically, from the point of view of a caller.

@@

\begin{figure}
{\scriptsize
\begin{code}
@import mergesort
@import swapsort

@@

Figure \ref{fig:parallelcode} shows the complete TSTL harness for
checking the parallel sorts.  This harness produces sequences of
integer pairs, where only the first integer is used in the comparison
operator of the sorting algorithms, and checks that the results of
{\tt sort\_parallel} calls are sorted..
Running TSTL's random tester without checking for determinism, we can
see that both sorting algorithms always produce sorted output.  The
horizontal determinism detector, however, always generates a 
test case showing the {\tt swapsort} behaving nondeterministically in
less than 3 seconds, thus
answering yes to {\bf  RQ1}.  If we comment out the call to {\tt
  swap\_sort\_parallel} and only check the parallel merge sort, however, TSTL
finds no nondeterministic behavior.  Running tests of the merge sort
with nondeterminism detection turned on results in a slowdown of approximately
40\% ({\bf RQ2} with little variance.

@@

{\scriptsize
\begin{code}
data0 = []    
val0 = 3      
val1 = -9    
data0.append((val0,val1)) 
val2 = -1                         
data0.append((val1,val2))
data0.append((val1,val1))  
sorted0 = swapsort.swap\_sort\_parallel(data0)      
\end{code}
}

%For this simple example, more sophisticated reduction strategies
%are not needed; we only want to see the code behaving
%nondeterministically, and do not care about preserving a higher probability of nondeterminism.

@@

\begin{figure}
\centering 
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisddmin}
\caption{Standard delta-debugging (38s)}
\label{fig:r1}
\end{subfigure}
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisforcep}
\caption{P=0.5, 100 samples (2131s)}
\label{fig:r2}
\end{subfigure}
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisforceprep}
\caption{P=0.5, 10 samples, 10 replications (617s)}
\label{fig:r3}
\end{subfigure}
\caption{Delta-debugging of the same {\tt redis-py} nondeterministic test}
\end{figure}

The {\tt redis-py} \cite{redispy} module implements a (very) widely-used Python interface
to the popular Redis in-memory database, cache, and
message broker \cite{redis}.  The TSTL {\tt redis-py} harnesses can be found in
the {\tt examples/redis} directory in the TSTL GitHub repo \cite{tstl}.

Using TSTL's harness for {\tt redis-py}, both {\tt redis-py} and Redis
itself can be tested; unfortunately, generating stand-alone
high-coverage regression tests for {\tt redis-py} and Redis using TSTL
has proven difficult, as numerous Redis commands introduce nondeterministic
behavior:  thus the resulting tests are very often \emph{flaky}.
Using AFL, we can see that {\tt redis-py} has a 
stability of only 56.26\%, a clear indicator of significant nondeterminism. Some
of the problematic commands are obvious simply by inspection (e.g.,
{\tt randomkey, srandmember}).  And, given an understanding of Redis
semantics, it is also clear that the various
commands producing data with a limited lifetime (e.g., {\tt expire,
  pexpire}) introduce timing-based nondeterminism.

However, that a command such as {\tt restore} takes an expiration
argument is not obvious to a non-expert, nor is the behavior of {\tt
  spop}, which pops a random value from a set.  Moreover, it is
difficult to guess whether the {\tt pipe} mechanism, which allows a
large sequence of commands to be queued up in a pipe and executed all
at once, introduced the potential for nondeterminism (does it execute
sequentially before other command are handled, or is it in parallel
with further commands).  Deep experience with Redis would make these
issues clear, but tests using a library are often written by those not
intimately familiar with the semantics of every library call (otherwise they would not have
been as likely to introduce flaky tests in the first place).  This is
even more true in the case of automated test generation, where the
test engineer is often chosen for expertise in test generation tools,
not in the domain of the software under test (SUT), and is seldom the
original developer of the code.

Using the original TSTL harness, just over 20\% of all {\tt redis-py}
regression tests (of length 200) generated were flaky.  Using the
nondeterminism checker to reduce flaky tests to a minimal set allows
us to simply set up an overnight run that will, in 12 hours, produce
a set of minimal nondeterministic tests exposing the root sources of
nondeterminism/flakiness: we set the harness to repeatedly run, then
produce a regression test.  If the test is flaky (exhibits
nondeterminism in result, over a set of 3 runs), we use the TSTL reducer to shrink it to a
normaliized \cite{onetest} test.  At the end of the testing period,
all distinct tests exhibiting nondeterminism are reported to the
user.  The process reduces the large number of different tests
exhibiting flakiness to a small set of tests, each short, showing a
different cause of nondeterminism.  The set produced is not perfect:
There is some duplication of causes, but it is sufficient for
identifying causes of nondeterminism.

@@

\begin{figure}[t]
{\scriptsize
\begin{code}
@import inference\_algs
@import datarray
<@
def flat\_sort(v):
    return (sorted(map(flat\_sort,v),key=repr) if type(v) in [list,tuple]
                else (flat\_sort(list(v.items())) if type(v) == dict else v))

@@

\{Exception\} print(flat\_sort(inference\_algs.calc\_marginals\_simple(<cpts>,
   <evidence>)))
\{Exception\} print(flat\_sort(inference\_algs.calc\_marginals\_sumproduct(<cpts>,
   <evidence>)))
\{Exception\} print(flat\_sort(inference\_algs.calc\_marginals\_jtree(<cpts>,
   <evidence>)))
\end{code}
}
\caption {Complete TSTL harness for finding the hash-order bug in the datarray
  inference algorithms.}
\label{hashbug}
\end{figure}

@@

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  The {\tt flat\_sort} function is
needed because we care about actual differences in probability values,
not simply the order of list, tuple, or dictionary items.  Running
this harness consistently requires less than 10 seconds to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm), answering {\bf RQ1} again in the affirmative.  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds ({\bf RQ3}) on average.  Interestingly, the way that {\tt
  python-afl} is used in TSTL means that AFL stability is 100\% even
for the original, broken, code,
because the {\tt PYTHONHASHSEED} has already been chosen before the
fork in AFL.  Removing the nondeterministic call, we can see that the cost of
checking for process nondeterminism, with no delay between operations,
is high, a mean 93\% slowdown ({\bf RQ2}.  This is due to the high cost of subprocess
creation and communication. However, even with this essentially worst
case behavior (where test runtime is very low compared to process overheads),
we sacrifice less than half of the test executions.

>>>>

Table \ref{tab:overviewexp} shows a summary of our experimental
results.  In all cases, our approach was able to detect interesting
nondeterminism (in fact, all nondeterminism of which we are aware).  The overhead on testing imposed by nondeterminism
checks varied considerably, but in the worst case did not double the
average cost of each action (the expected overhead, since we
essentially execute each test twice), and thus never halved the amount of
testing performed.  Test reduction always reduced the size of
generated tests by 85\% or more, in less than 15 minutes, thus showing that reduction
has a large payoff at an acceptable price.  Values are, unless otherwise noted, the mean of 30 runs.

@@

Our first subject is an implementation of a parallel merge
sort ({\url{https://gist.github.com/stephenmcd/39ded69946155930c347}). When a sorting
algorithm is implemented using concurrency, a key question to ask is
whether the ordering of (equal under comparison) elements is consistent.
This is related to the question of whether a sort is stable,
but imposes a less strict requirement; a concurrent sort
might not be stable, but might still produce the same (non-stable)
ordering from the same provided sequence, every time.  While a parallel
implementation of an algorithm will likely have internal
nondeterminism, it is often expected to
behave deterministically, from the point of view of a caller.

@@

\begin{comment}
\begin{figure}
{\scriptsize
\begin{code}
@import mergesort
@import swapsort

@@

Figure \ref{fig:parallelcode} shows the complete TSTL harness for
checking the parallel sorts.
\end{comment}

Our TSTL harness for parallel sorting produces sequences of
integer pairs, where only the first integer is used in the comparison
operator of the sorting algorithms, and checks that the results of
are sorted..
Running TSTL's random tester without checking for determinism, we can
see that both sorting algorithms always produce sorted output.  The
horizontal determinism detector, however, always generates a 
test case showing the {\tt swapsort} behaving nondeterministically in
less than 3 seconds, thus
answering yes to {\bf  RQ1}.  If we comment out the call to {\tt
  swap\_sort\_parallel} and only check the parallel merge sort, TSTL
finds no nondeterminism.  This results in a slowdown of approximately
40\% ({\bf RQ2}).

@@

\begin{comment}
{\scriptsize
\begin{code}
data0 = []    
val0 = 3      
val1 = -9    
data0.append((val0,val1)) 
val2 = -1                         
data0.append((val1,val2))
data0.append((val1,val1))  
sorted0 = swapsort.swap\_sort\_parallel(data0)      
\end{code}
}
\end{comment}
%For this simple example, more sophisticated reduction strategies
%are not needed; we only want to see the code behaving
%nondeterministically, and do not care about preserving a higher probability of nondeterminism.

@@

\begin{figure}
\centering 
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisddmin}
\caption{No modification}
\label{fig:r1}
\end{subfigure}
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisforcep}
\caption{N=100 (2131s)}
\label{fig:r2}
\end{subfigure}
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisforceprep}
\caption{M=10, N=10}
\label{fig:r3}
\end{subfigure}
\caption{Reducing a {\tt redis-py} nondeterministic test}
\end{figure}

The {\tt redis-py} \cite{redispy} module implements a (very) widely-used Python interface
to the popular Redis in-memory database, cache, and
message broker \cite{redis}.  
Using TSTL's harness for {\tt redis-py}, both {\tt redis-py} and Redis
itself can be tested; unfortunately, generating stand-alone
high-coverage regression tests for them
has proven difficult, as numerous Redis commands introduce nondeterministic
behavior:  thus the resulting tests are very often \emph{flaky}.
Using AFL, we can see that {\tt redis-py} has a 
stability of only 56.26\%, a clear indicator of significant nondeterminism. Some
of the problematic commands are obvious simply by inspection (e.g.,
{\tt randomkey, srandmember}). It is also clear that 
commands producing data with a limited lifetime (e.g., {\tt expire,
  pexpire}) introduce timing-based nondeterminism.

However, that a command such as {\tt restore} takes an expiration
argument is not obvious to a non-expert, nor is the behavior of {\tt
  spop}, which pops a random value from a set.  Moreover, it is
difficult to guess whether the {\tt pipe} mechanism, which allows a
large sequence of commands to be queued up and executed
at once, introduces the potential for nondeterminism. Deep experience with Redis would make these
issues clear, but tests using a library are often written by those not
intimately familiar with the semantics of every library call (otherwise they would not have
been as likely to introduce flaky tests in the first place).  This is
even more true in the case of automated test generation, where the
test engineer is often chosen for expertise in test generation tools,
not in the domain of the SUT, and is seldom the
original developer.

Using the original TSTL harness, just over 20\% of all {\tt redis-py}
regression tests (of length 200) generated were flaky.  Using the
nondeterminism checker to reduce flaky tests to a minimal set allows
us to set up an overnight run that will, in 12 hours, produce
a set of minimal nondeterministic tests exposing the root sources of
nondeterminism/flakiness: we set the harness to repeatedly run, then
produce a regression test.  If the test is flaky (exhibits
nondeterminism in result, over a set of 3 runs), we use the TSTL reducer to shrink it to a
normaliized \cite{onetest} test.  At the end of the testing period,
all distinct tests exhibiting nondeterminism are reported to the
user.  The process reduces the large number of different tests
exhibiting flakiness to a small set of tests, each short, showing a
different cause of nondeterminism.  
There is some duplication of causes, but the tests are sufficient for
identifying all nondeterminism sources.

@@

\begin{comment}
\begin{figure}[t]
{\scriptsize
\begin{code}
@import inference\_algs
@import datarray
<@
def flat\_sort(v):
    return (sorted(map(flat\_sort,v),key=repr) if type(v) in [list,tuple]
                else (flat\_sort(list(v.items())) if type(v) == dict else v))

@@

\{Exception\} print(flat\_sort(inference\_algs.calc\_marginals\_simple(<cpts>,
   <evidence>)))
\{Exception\} print(flat\_sort(inference\_algs.calc\_marginals\_sumproduct(<cpts>,
   <evidence>)))
\{Exception\} print(flat\_sort(inference\_algs.calc\_marginals\_jtree(<cpts>,
   <evidence>)))
\end{code}
}
\caption {Complete TSTL harness for finding the hash-order bug in the datarray
  inference algorithms.}
\label{hashbug}
\end{figure}
\end{comment}

@@

\begin{comment}
Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  The {\tt flat\_sort} function is
needed because we care about actual differences in probability values,
not simply the order of list, tuple, or dictionary items.
\end{comment}
Running our TSTL
harness to test {\tt datarray} consistently requires less than 10 seconds to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm), answering {\bf RQ1} again in the affirmative.  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds ({\bf RQ3}) on average.  Interestingly, the way that {\tt
  python-afl} is used in TSTL means that AFL stability is 100\% even
for the original, broken, code,
because the {\tt PYTHONHASHSEED} has already been chosen before the
fork in AFL.  Removing the nondeterministic call, we can see that the cost of
checking for process nondeterminism, with no delay between operations,
is high, a mean 93\% slowdown ({\bf RQ2}.  This is due to the high cost of subprocess
creation and communication. However, even with this essentially worst
case behavior (where test runtime is very low compared to process overheads),
we sacrifice less than half of the test executions.

### 2020-01-16 ISSTA/implementation.tex
ok

We provide support for the real-world sources of
nondeterminism described in Section \ref{sec:sources}, via
control over timing and/or execution in a new process.  Changing the timing of
steps is also effective in exposing concurrency-related
nondeterminism, since it often changes the
interleaving of threads.

Because TSTL has an interface to AFL \cite{aflfuzz}, we can use
AFL's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism.  Moreover, since
there is substantial overhead in nondeterminism detection, AFL can be
used to predict whether tested Python code is a good candidate for
checking for horizontal nondeterminism; if the
AFL \emph{stability} statistic (see the AFL documentation for details
\cite{aflfuzz}) is lower than 100\%, it may indicate horizontal
nondeterminism.
Because of the idiosyncrasies of AFL instrumentation and process
behavior, this is not always a reliable guide, but it is a very low
cost indicator produced as a by-product of fuzzing for other problems.

>>>>

We provide support for the real-world sources of
nondeterminism described in Section \ref{sec:sources}, via
control over timing and/or execution in a new process.  Changing the timing of
steps is also often effective in exposing concurrency-related
nondeterminism.

Because TSTL has an interface to AFL \cite{aflfuzz}, we can use
AFL's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism.  Moreover, since
there is substantial overhead in nondeterminism detection, AFL can be
used to predict whether tested Python code is a good candidate for
checking for horizontal nondeterminism; if the
AFL \emph{stability} statistic \cite{aflfuzz} is lower than 100\%, it may indicate horizontal
nondeterminism.
Because of the idiosyncrasies of AFL instrumentation and process
behavior, this is not always a reliable guide, but it is a very low
cost indicator.
